Inferring ecological interactions

### 2017-02-03 ms/ms_interactions.tex
Change margin, add line numbers and double space manuscript
Worked on the Introduction

\begin{document}

@@

Biotic interactions form the backbone of biodiversity
\cite{bascompte_plant-animal_2007}. Understanding the strength, direction, and
symmetry of species interactions provides insights into the maintenance of
assemblages and potential threats from anthropogenic change. Combined with
abiotic factors, biotic interactions constrain the distributions of species,
shape the evolution of phenotypes, and influence the stability of natural
systems \cite{schleuning_predicting_2015}. A common tool for understanding
interaction are ecological networks, in which each species (nodes) are connected
based on the strength and direction of interactions (links)(Jordano 1987).
Documenting and analyzing these networks is essential to understand the rules
underlying pairwise species interactions (Bascompte et al. 2006), quantifying
ecosystem functioning (Thomson et al 2012), and predicting the biodiversity
consequences due to anthropogenic change (ref).

Species traits are at the core of biological interactions (Bartomeus et al.
2016). In this paper, we defined traits as phenotypic adaptations to exploit abiotic or
biotic resources. We used traits as our principal mechanism to explain when
species can interact. The foremost requirement for species interactions is
species co-occurrence, which is itself driven by traits through environmental
filtering and local adaptation (Hillerislambers, Mayfield, but see Kraft). If
species co-occur, they may interact if their respective traits promote the
efficiency of resource extraction. In addition, traits can be represented by
phylogenetic information, in which evolution history constrains trait evolution
either directly through phylogenetic conservatism or indirectly through exposure
to similiar environmental conditions (Graham and Wiens 2005).

The traits promoting species co-occurrence and interaction may be hidden from
us, either through lack of sampling, or due to their complexity. In these cases,
the co-occurrence of species is used as a proxy for interaction. Given the
correlation in distribution or abundance of species across space or time,
co-occurrence based methods attempt to circumvent trait-based models by assuming
that species which consistently co-occur, and whose populations co-vary, may
have important biotic interactions. We therefore see co-occurrence based methods
as an extension of specific trait-based models when the mechanisms promoting
interactions are unknown or unobserved.

In spite of increasing efforts, it is unlikely that ecologists will empirically
document the enormous number of interactions existing in nature
(Morales-Castilla et al. 2015). Therefore, complementing field work with models
of species interactions is a promising avenue for understanding the complexity
of the natural world(ref). While direct inference will often be preferable, many
interactions are rare, infrequent, or impossible to detect by human observation.
Model-based inference for ecological networks provides three important aspects,
i) it establishes mechanistic relationships among species, ii) it estimates
uncertainty given empirical observation, and iii) it provides an avenue to
predict future interactions. We follow Poisot (2015) and Cazelles (2016) in
formulating our general conceptual model of species interactions (Fig. 1). There
is a pool of potential interactions, termed a metaweb, in which species could
interact based on co-occurrence. Any given observation we make is an attempt to
recover these relationships using empirical data. We use these data to fit a
model based on either species taxonomic identity or phenotypic characteristics.
This model can used to predict the interactions among observed species, given
the uncertainty due to spatial and temporal stochasticity of interactions, or a
new network, based on the presence of a novel species in the assemblage. From
these predictions, we can estimate structural network properties, such as
connectance and nestedness, while simultaneously propagating our uncertainty
throughout the analysis. In this paper, we review the methods proposed to build
models of species interactions, emphasizing each method's utility for predicting
the position of a novel species in the network.

From this conceptual framework, we can decide on a set of optimal criteria to
compare existing models. While there is unlikely to be any single method that
meets all our expectations, by defining a set of standards, we can more
transparently judge the relative merits of each method for our particular goal.
Our foremost goal is to generate a prediction of the links among co-occurring
species based on an estimated probability of interaction. This prediction should
come with a statement of uncertainty regarding our confidence in parameter
estimates based on empirical data. An ideal method will make reasonable
assumptions about the interdependence of observations, and be relatively
insensitive to low sampling effort. The mechanism determining the links among
species should be clear and grounded within ecological theory, such that we can
use our method to make a prediction about the placement of a new species in the
network based on a fitted model.

@@

>>>>

\begin{document}
% Add line number
\linenumbers
\modulolinenumbers[1]

% Double space
\doublespacing

@@

Biotic interactions form the backbone of biodiversity
\cite{bascompte_plant-animal_2007}. Understanding the strength, direction, and
symmetry of species interactions provides insights into the maintenance of
assemblages and potential threats from anthropogenic change. Combined with
abiotic factors, biotic interactions constrain the distributions of species,
shape the evolution of phenotypes, and influence the stability of natural
systems \cite{schleuning_predicting_2015}. A common tool for understanding
interaction are ecological networks, in which each species (nodes) are connected
based on the strength and direction of interactions (links)(Jordano 1987).
Documenting and analyzing these networks is essential to better understand the
rules underlying pairwise species interactions (Bascompte et al. 2006),
quantifying ecosystem functioning (Thomson et al 2012), and predicting the
biodiversity consequences due to anthropogenic change (ref).

The foremost requirement for species interactions is species co-occurrence,
which is itself driven by species traits through environmental filtering and
local adaptation (Hillerislambers, Mayfield, but see Kraft). Knowing where
species are located gives us insights into the interactions they may have, but
to more deeply understand the mechanisms that influence the relationships among
species, traits are necessary. As such, species traits are at the core of
biological interactions (Bartomeus et al. 2016). Here, we defined traits as
phenotypic or genotypic adaptations, that may vary through time, to exploit
abiotic or biotic resources. Note that this definition indirectly include
phylogeny as a form of trait. From this definition, conceptually, if species
co-occur, we assume that they may interact only if their respective traits
promote the efficiency of resource extraction. When traits are represented by
phylogenetic information, it is the evolution history that constrains trait
evolution either directly through phylogenetic conservatism or indirectly
through exposure to similiar environmental conditions (Graham and Wiens 2005).

In practice, the traits promoting species co-occurrence and interaction may be
hidden from us, either because they were not sampled or due to their complexity.
In such cases, co-occurrence of species is used as a proxy for interaction.
Given the correlations in occurrence or abundance of species across space or
through time, co-occurrence based methods attempt to circumvent trait-based
models by assuming species that consistently co-occur, and whose populations
co-vary, may have important biotic interactions. In this context, co-occurrence
based methods are used to describe the potential relationships species may have
while trait-based models consider the mechanisms promoting interactions.

In spite of increasing efforts, it is unlikely that ecologists will empirically
document the enormous number of interactions existing in nature
(Morales-Castilla et al. 2015). Therefore, complementing field work with models
of species interactions is a promising avenue for understanding the complexity
of the natural world(ref). While direct inference is often preferable, many
interactions are rare, infrequent, or impossible to detect through human
observation. Model-based inference for ecological networks provides three
important aspects to characterize species interactions, i) it establishes
mechanistic relationships among species, ii) it estimates uncertainty given
empirical observation, and iii) it provides an avenue to predict future
interactions. We follow Poisot (2015) and Cazelles (2016) in formulating our
general conceptual model of species interactions (Fig. 1). There is a pool of
true interactions, a metaweb, in which individuals of different species co-occur
resulting in interactions among species. Any given observation we make on this
metaweb is an attempt to recover the relationships among species using empirical
data. Based on these data, model(s) are built to characterize and elucidate the
relationships among species. Some models rely exclusively co-occurrence of
individuals of different species at particular locations and may also include
information on phylogenetic history and species traits. These models can than be
used to predict the interactions among observed species, given an uncertainty
due to the spatial and/or temporal stochasticity of the interactions. With these
models it is also possible to define new network, based on the presence of a
novel species in the assemblage. From these predictions, we can estimate
structural network properties, such as connectance and nestedness, while
simultaneously propagating our uncertainty throughout the analysis. In this
paper, we reviewed the methods proposed to build models of species interactions,
emphasizing each method's utility for predicting the position of a novel species
in the network.

From this conceptual framework, we can decide on a set of optimal criteria to
compare existing approaches. While it is unlikely that any single method meets all
our expectations, by defining a set of standards, we can more transparently
judge the relative merits of each method to reach particular goals. The most
important aspect of a method is for it to generate accurate prediction of the
links among co-occurring species based on an estimated probability of
interaction. This prediction should come with a statement of uncertainty
regarding our confidence in parameter estimates based on empirical data. An
ideal method would also be able to make reasonable assumptions about the interdependence of
observations, and be insensitive to low sampling effort. The
mechanism determining the links among species should be clear and grounded
in ecological theory such that we can use our method to make a prediction
about the placement of a new species in the network based on a fitted model.

@@

A handful of soil is the home of countless number of organisms. Soils host about
the quarter of Earth biodiversity (Decaëns et al 2006) and is recognised as the
third biotic frontier (André et al 1994). But three crucial points arise when
dealing with soil biodiversity and soil biotic interactions, (i) it is hard to
observe organisms without destroying their habitat, (ii) most of soil organisms
are tiny and cannot be bred or manipulated and (iii) soil organisms are prone to
a “Linnean deficit”. Biotic interactions in soils are shaped by soil
specificities that lead to an intense entanglement of trophic and non-trophic
interactions. For instance, soils are physically structured by the activity of
ecosystem engineers (earthworms, termites, ants, roots) that modifies, by their
own niche construction, the accessibility to resources for other soil organisms
(Jones et al 1994).

Body size of soil organisms ranges from nanometric scale (phage, bacteria) to
decimetric one (earthworms, moles, etc), leading to nested home ranges. Soils
are highly organized environments (in the three dimensions and in time). This
permits high level of niche segregation (to diminish competition) and numerous
way to escape from predators. On the most famous example of soil trophic network
comes from short grass meadows in Colorado (Hunt et al. 1987), in which nodes
were a mix of taxa, feeding guilds and pools of organic matter. The observation
of soil species interactions is hard and interaction can be inferred by various
ways. In the case of trophic interaction, this can be achieved through, for
example, gut content analyses (observation-based or DNA-based), (bio)chemical
analyses (eg PLFA, stable isotopes content), predator-prey body size
relationship, co-occurrence analysis, machine learning or a combination of them
(Digelet al 2014). Given the lack of knowledge on microbes’ biology and ecology,
co-occurrence is often interpreted as interaction (Barberan et al 2012).
Conversely to the green-world hypothesis (predator limitation of herbivores can
cascade to facilitate plant growth), soil food webs are assumed to be bottom-up
controlled detritus-based communities, but yet the ecosystem role of predators
in detrital food webs is less explored. Models of energy fluxes through soil
trophic networks remains rare (de Vries et al 2013), and nutrient cycling is
generally ignored while connecting above and belowground trophic networks (Zou
et al 2016). Hence, soil food webs studies are mainly descriptive and have poor
predictive ability.

Given the importance of plant for humankind, it is crucial to understand the how
multiple interactions control plant production for food, fibre and wood. Soils
are the place of numerous symbiotic relationships between plants and bacteria
(eg Rhizobia) or fungi (mycorrhiza). The soil microbial loop an example of
hypothesis that links multiple partners of the rhizospheric microbial food web
to explain protozoa (predators) effects on plant growth. As carbon-rich root
exudates (high C to N ratio) are continuously released into the plant
rhizosphere, those soil bacteria that consume the exudates need other sources of
nitrogen to support their growth. It is suggested that the bacteria therefore
increase their assimilation of indigenous soil organic nitrogen. When protozoa
subsequently consume the bacteria, part of this nitrogen will be released into
the soil inorganic nitrogen pool, which will increase the nitrogen pool
available for plant uptake. It has also been demonstrated that signal molecules
emitted by invertebrates can mediate plant growth and plant disease immunity
(resistance or recovery). Interactions between soil organisms and roots also
shape the plant community assembly and the aboveground trophic interactions, up
to fourth levels (plant-herbivore-parasite-parasitoïd).

Recently, DNA-based approach has been democratised to study co-occurrence of
cryptic interactions like Ericaceous plant-fungus symbioses (Toju et al 2016),
earthworm-gut bacteria symbioses (Moller et al 2015) or predator-prey
interactions (Bohan xxx). The tremendous development of next-generation DNA
sequencing will surely provide the most comprehensive and accurate measure of
trophic links in ecosystems, regardless of the size of organisms and without a
priori on the species consumed by a given organism, by matching sequences
amplified from the gut content to wide gene databases (Pompanon et al. 2012).
Finally, the integration of biogeography and network theory will yield important
results and further our understanding and forecasting ability of species
interactions, in a context of global changes.
